---
title: "PredictiveModel"
author: "Jenny Rogers"
date: "June 15, 2020"
output: html_document
---


```{r}

library(tidyverse)
library(lubridate)

```



this is a script to predict the probability of occurrence of different life stages of typha due to changing hydraulics in the Los Angeles River. The variables that we are using to measure habitat suitability inclue depth (germination/seedling survial, adult patches), velocity (adult patches), and temperature (germination). 

In this first chunk of code, we read in and tidy the projected hydraulic data from HEC-RAS simulations that model hydraulics under different waste water reuse. 
We then read in and tidy the temperature data


```{r}

#this is example data for node F57C
hydlics <- read.csv("hecras.csv")
hydlics <- hydlics %>%
  select(1: 3, contains("Depth"), contains("Vel")) %>% 
  mutate(Vel_LOB = Avg..Vel...ft.s..LOB * 0.3048, #convert f/s to m/s
         Depth_LOB = Hydr..Depth..ft..LOB * 30.48, #convert ft to cm
         Vel_MC  = Avg..Vel...ft.s..MC * 0.3048, 
         Depth_MC  = Hydr..Depth..ft..MC * 30.48,
         Vel_ROB  = Avg..Vel...ft.s..ROB * 0.3048, 
         Depth_ROB  = Hydr..Depth..ft..ROB * 30.48) %>% 
  select(-(4:9)) %>% 
  mutate(DateTime = mdy_hm(DateTime),
         year = year(DateTime),
         month = month(DateTime),
         day = day(DateTime)) %>% 
  pivot_longer(cols = 4:9, names_to = "variable", values_to = "value") %>% 
  separate(variable, into = c("variable", "bank"), sep = "_") %>% 
  pivot_wider(names_from = variable, values_from = value) %>% 
  rename(vel_m_s = Vel,
         depth_cm = Depth) %>% 
  mutate(species = "typha latifolia") #add in a species name to the germination model


#this is example temperature data from the observed Temp data for compton creek
temperature <- read.csv("temperature_data.txt")
temperature <- temperature %>% 
  separate("yyyymmdd.Hr.Min.Sec.Temp", into = c("date", "time", "temperature"), sep = ",") %>% 
  unite(col = date, c(date, time), sep = ":") %>% 
  mutate(DateTime = ymd_hms(date),
         year = year(DateTime),
         month = month(DateTime, label = TRUE),
         day = day(DateTime),
         temperature = as.numeric(temperature)) %>% 
  select(-date) %>% 
  group_by(month, day) %>% 
  summarise(high_temperature_c = max(temperature),
            temperature_range_c = max(temperature)- min(temperature))

```


In this second chunk of code, we read in the five models used to predict different outcomes of habitat suitabilty 

```{r}

#velocity model for typha patches

#logistic regression model, so the prediction will be log odds and will need to be transformed to be probablity of occurrence
load("vel_ptch_mdl.rda")
summary(vel_ptch_mdl)



#temperature models for typha germination

#linear model with a quadratic term for percent germination
load("tmp_germ_mdl.rda")
summary(tmp_germ_mdl)



#depth models for typha germination, seedling survival, and patches

#linear model with a quadatic term for germinaion percent
load("dep_germ_mdl.rda")
summary(dep_germ_mdl)

#linear model with a quadratic term for seedling survial percent
load("dep_sdlng_mdl.rda")
summary(dep_sdlng_mdl)

#linear model with a quadratic term for occurrence
load("dep_ptch_mdl.rda")
summary(dep_ptch_mdl)


```


in this chunk of code, we apply the models to the projected stream hydraulics
```{r}

#adult probability of occurrence velocity
hydlics$pred_ptch_vel <- predict(vel_ptch_mdl, newdata = hydlics) #predict with new data
hydlics$prob_ptch_vel <- exp(hydlics$pred_ptch_vel)/(1+exp(hydlics$pred_ptch_vel)) #convert log odds to prob
hydlics <- hydlics %>% select(-pred_ptch_vel) #remove the log odds column


#adult occurrence depth
hydlics$pred_ptch_dep <- predict(dep_ptch_mdl, newdata = hydlics) #occurrence


#germination depth
hydlics$pred_germ_dep <- predict(dep_germ_mdl, newdata = hydlics) #germination percent

#germination temperature
temperature$pred_germ_tmp <- predict(tmp_germ_mdl, newdata = temperature)


#seedling depth
hydlics$pred_sdlng_dep <- predict(dep_sdlng_mdl, newdata = hydlics) #seedling survival percent


```


In this chunk of code well plot
```{r}

#physicla metrics
ggplot(data = hydlics, mapping = aes(x = DateTime, y = vel_m_s))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Node F57C Velocity (m/s)")

ggplot(data = hydlics, mapping = aes(x = DateTime, y = depth_cm))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Node F57C Depth (cm)")

ggplot(data = temperature, mapping = aes(x = day, y = high_temperature_c))+
  geom_line()+
  facet_wrap(vars(month))+
  labs(x = "Day", y = "Max Temperature (C)")

ggplot(data = temperature, mapping = aes(x = day, y = temperature_range_c))+
  geom_line()+
  facet_wrap(vars(month))+
  labs(x = "Day", y = "Temperature Range (C)")





#predicted occurrence 
ggplot(data = hydlics, mapping = aes(x = DateTime, y = prob_ptch_vel))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Prob of Typha Occurrence due to Velocity")

ggplot(data = hydlics, mapping = aes(x = DateTime, y = pred_ptch_dep))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Prob of Typha Occurrence due to Depth")

ggplot(data = hydlics, mapping = aes(x = DateTime, y = pred_germ_dep))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Germination (%)")

ggplot(data = temperature, mapping = aes(x = day, y = pred_germ_tmp))+
  geom_line()+
  facet_wrap(vars(month))+
  labs(x = "Day", y = "Germination (%)")

ggplot(data = hydlics, mapping = aes(x = DateTime, y = pred_sdlng_dep))+
  geom_line()+
  facet_wrap(vars(bank))+
  labs(x = "Date", y = "Seedling Survival (%)")





```



In this section, we experiment with ways to 
```{r}

synthesis <- hydlics %>% 
  mutate(adult = ifelse(prob_ptch_vel <= pred_ptch_dep, prob_ptch_vel, pred_ptch_dep)) %>% 
  group_by(bank, year, month) %>% 
  summarise(min_ptch = min(adult),
            average_ptch = mean(adult),
            min_germ = min(pred_germ_dep),
            average_germ = mean(pred_germ_dep)) %>% 
  ungroup() 

ggplot(data = synthesis, mapping = aes(x = month, y = min_ptch))+
  geom_line()+
  labs(x = "Month", y = "Min Probability") +
  facet_grid(rows = vars(as.character(year)), cols = vars(as.character(bank)))



a <- ggplot(data = synthesis, mapping = aes(x = month, y = average_ptch))+
  geom_line()+
  labs(x = "Month", y = "Average Probability") +
  facet_grid(rows = vars(as.character(year)), cols = vars(as.character(bank)))+
  theme(text = element_text(size=40), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
ggsave(a, filename = "average.tiff", width = 20, height = 25, dpi = 300, compression = "lzw")



Germination <- synthesis %>% 
  filter(month %in% c(5:8))

a <- ggplot(data = Germination, mapping = aes(x = month, y = average_germ))+
  geom_line()+
  labs(x = "Month", y = "Germination (%)") +
  facet_grid(rows = vars(as.character(year)), cols = vars(as.character(bank)))+
  theme(text = element_text(size=40), panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
        panel.background = element_blank(), panel.border = element_rect(colour = "black", fill=NA, size=.5))
ggsave(a, filename = "average.tiff", width = 20, height = 25, dpi = 300, compression = "lzw")






most_consecutive_val = function(vector, val = 1) { 
  with(rle(vector), max(lengths[values == val]))
}

grm <- hydlics  %>% 
    group_by(bank, year) %>% 
    summarize(germ_depth_suitability = most_consecutive_val(vector=abs(diff(which(pred_germ_dep >= 35))==1)))
write.csv(grm, file = "suitability.csv")

suitability <- 
  data.frame(
    "year" = grm$year, 
    "bank" = grm$bank,
    "lifestage" = "germination", 
    "variable" = "depth", 
    "suitability" = ifelse(grm$germ_depth_suitability >= 336, "suitable", "unsuitable")) 
suitability <- suitability %>% unique()
write.csv(suitability, file = "suitability.csv")

```
